Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-09-19 paper/paper.Rmd
add example and plot for `"cv"` and update description

```{r buffer-non-eval, eval=FALSE}
library("mlr3")
library("mlr3spatiotempcv")
task = tsk("ecuador")
rsmp_buffer = rsmp("spcv_buffer", theRange = 1000)
rsmp_buffer$instantiate(task)

@@

```{r buffer-eval, fig.cap='Visualization of the spatial buffering method from package blockCV (method "spcv\\_buffer" in  mlr3spatiotempcv). The buffer distance is 1000 m.', out.width="40%", fig.pos="ht",echo=FALSE}
library("mlr3")
library("mlr3spatiotempcv")
task = tsk("ecuador")
rsmp_buffer = rsmp("spcv_buffer", theRange = 1000)
rsmp_buffer$instantiate(task)

@@

```{r disc-non-eval, eval=FALSE}
rsmp_disc = rsmp("spcv_disc", folds = 100, radius = 300L, buffer = 400L)
rsmp_disc$instantiate(task)

@@

```{r disc-eval, fig.cap='Visualization of one training set / test set combination generated with the leave-one-disc-out method from package sperrorest (method "spcv\\_disc" in mlr3spatiotempcv). The disc has a radius of 300 m and is surrounded by a 400-m buffer.', out.width="40%", echo=FALSE,fig.pos="hb"}
library("mlr3")
library("mlr3spatiotempcv")
task = tsk("ecuador")
rsmp_disc = rsmp("spcv_disc", folds = 100, radius = 300L, buffer = 400L)
rsmp_disc$instantiate(task)

@@

```{r coords-non-eval, eval=FALSE}
rsmp_coords = rsmp("spcv_coords", folds = 5)$instantiate(task)

@@

```{r coords-eval, fig.cap='Leave-one-block-out CV based on $k$-means clustering of the coordinates as implemented in package sperrorest (method "spcv\\_coords" in mlr3spatiotempcv).', out.width="40%", fig.pos="ht", echo=FALSE}
rsmp_coords = rsmp("spcv_coords", folds = 5)$instantiate(task)

@@

```{r custom-cv, fig.cap='Leave-one-level-out (custom) resampling from package mlr3 (method "custom\\_cv"). A factor variable is used to define the partitions. Each color represents one partition.', out.width="40%", fig.pos="ht", echo=FALSE}
breaks = quantile(task$data()$dem, seq(0, 1, length = 6))
zclass = cut(task$data()$dem, breaks, include.lowest = TRUE)

@@

Group-level CV is based on the idea that an inherent grouping structure is present the observations or the sampling design in space and/or in time, which can be used to determine partitions.
In contrast to blocking or clustering, the spatial or temporal location is not used explicitly, but additional metadata is exploited such as the membership in a specific agricultural management unit [@brenning2006], or in a site in hierarchical sampling [@kasurak2011].
<!--
This has also been referred to as "target-oriented validation" [@cast].
FIXME:
AB: I don't think I agree with this. My interpretation of the Meyer paper was that in the Meyer paper, locations or time points are the targets. Fields consist of many grid cells, at multiple location.
-->
Even though flexibility with respect to fold creation is limited due to the inherited grouping structure within the data, some implementations allow to specify the resulting number of folds by omitting parts of the dataset.

\pkg{mlr3spatiotempcv} supports two "grouping" implementations: (1) The "leave-location-and-time-out" family from package \pkg{CAST} and (2) "custom resampling" from the \pkg{mlr3} package.

@@

plot_env_single = autoplot(rsmp_env, size = 0.8, fold_id = 1, task = task)

plot_env_multi = autoplot(rsmp_env_multi, size = 0.8, fold_id = 1, task = task)

@@

```{r env-1-eval, fig.cap='Environmental leave-one-block-out CV from package blockCV using one (left, "distdeforest") and two (right, "distdeforest" and "slope") predictors to define blocks in the feature space.', out.width="100%", fig.pos="ht", echo=FALSE}
rsmp_env = rsmp("spcv_env",
  features = "distdeforest", folds = 5)

@@

plot_env_single = autoplot(rsmp_env,
  size = 0.6, fold_id = 1, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02))

plot_env_multi = autoplot(rsmp_env_multi,
  size = 0.6, fold_id = 1, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02))

library("patchwork")
plot_env_single + plot_env_multi + plot_layout(guides = 'collect')
```

@@

```{r llto, echo=FALSE, out.width="70%", fig.cap='Perspective plot of leave-location-and-time-out CV from package CAST (method "sptcv\\_cstf" and arguments time\\_var = "Date" and space\\_var = "SOURCEID"). The grey points are excluded from both the training and the test set in this fold.', fig.pos="ht"}
knitr::include_graphics("png/llto.png")
```

@@

```{r create-task}
b = mlr3::as_data_backend(ecuador)
task = TaskClassifST$new(
  id = "ecuador", b, target = "slides", positive = "TRUE",
  extra_args = list(
    coordinate_names = c("x", "y"), coords_as_features = FALSE,
    crs = "EPSG:32717")
)
```

@@

```{r rr-nspcv-1, cache.lazy = FALSE, results='hide'}
rsmp_nsp = rsmp("repeated_cv", folds = 4, repeats = 2)
rsmp_nsp$instantiate(task)
rr_nsp = resample(
  task = task, learner = learner,
  resampling = rsmp_nsp
)
```

@@

```{r rr-spcv-1, cache.lazy = FALSE, results='hide'}
rsmp_sp = rsmp("repeated_spcv_coords", folds = 4, repeats = 2)
rsmp_sp$instantiate(task)
rr_sp = resample(
  task = task, learner = learner,
  resampling = rsmp_sp
)
```

>>>>

```{r buffer-non-eval, eval=FALSE}
library("mlr3")
library("mlr3spatiotempcv")
task = tsk("ecuador")
rsmp_buffer = rsmp("spcv_buffer", theRange = 1000)

@@

```{r buffer-eval, fig.cap='Visualization of the spatial buffering method from package blockCV (method "spcv\\_buffer" in  mlr3spatiotempcv). The buffer distance is 1000 m.', out.width="40%", fig.pos="ht",echo=FALSE}
library("mlr3")
library("mlr3spatiotempcv")
task = tsk("ecuador")
rsmp_buffer = rsmp("spcv_buffer", theRange = 1000)
rsmp_buffer

@@

```{r disc-non-eval, eval=FALSE}
rsmp_disc = rsmp("spcv_disc", folds = 100, radius = 300L, buffer = 400L)
rsmp_disc

@@

```{r disc-eval, fig.cap='Visualization of one training set / test set combination generated with the leave-one-disc-out method from package sperrorest (method "spcv\\_disc" in mlr3spatiotempcv). The disc has a radius of 300 m and is surrounded by a 400-m buffer.', out.width="40%", echo=FALSE,fig.pos="ht"}
library("mlr3")
library("mlr3spatiotempcv")
task = tsk("ecuador")
rsmp_disc = rsmp("spcv_disc", folds = 100, radius = 300L, buffer = 400L)
rsmp_disc

@@

```{r coords-non-eval, eval=FALSE}
rsmp_coords = rsmp("spcv_coords", folds = 5)

@@

```{r coords-eval, fig.cap='Leave-one-block-out CV based on $k$-means clustering of the coordinates as implemented in package sperrorest (method "spcv\\_coords" in mlr3spatiotempcv).', out.width="40%", fig.pos="ht", echo=FALSE}
rsmp_coords = rsmp("spcv_coords", folds = 5)

@@

```{r custom-cv, fig.cap='Leave-one-level-out (custom) resampling from package mlr3 (method "custom\\_cv"). A factor variable is used to define all partitions.', out.width="40%", fig.pos="ht", echo=FALSE}
breaks = quantile(task$data()$dem, seq(0, 1, length = 6))
zclass = cut(task$data()$dem, breaks, include.lowest = TRUE)

@@

The `"cv"` resampling strategy performs random, non-spatial partitioning.
To use it in a spatiotemporal context, one can set the "group" column role in a mlr3 Task object for specific variable(s).
This tells the resampling strategy to honor these grouping(s) during partitioning.
In the final folds, these groups of observations will either belong to the test or training set and by this create possibly spatiotemporal resampling folds.

The idea is to make use of inherent grouping structure in the observations or the sampling design in space and/or in time to determine partitions.
In contrast to geometric blocking or clustering, the spatial or temporal location is not used explicitly, but additional metadata is exploited such as the membership in a specific agricultural management unit [@brenning2006], or in a site in hierarchical sampling [@kasurak2011].
In \pkg{mlr} this concept was called "blocking".

In the following example we create a custom grouping of eight classes using a kmeans clustering.
Note that in practice usually the group creation would not be done in an unsupervised way but by making use of an inherent grouping information from the dataset.
To showcase the honoring of the groups a uneven amount of folds (with respect to the number of groups) is requested.
The test set in the first and second fold is composed out of three groups while the third fold is only made up from two groups.

```{r cv-non-eval-1, eval=FALSE}
task_cv = tsk("ecuador")
group = as.factor(kmeans(task$coordinates(), 8)$cluster)
task_cv$cbind(data.frame("group" = group))
task_cv$set_col_roles("group", roles = "group")

rsmp_cv_group = rsmp("cv", folds = 3)$instantiate(task_cv)

print(rsmp_cv_group$instance)
```

```{r cv-non-eval-2, fig.cap='Cross-Validation at the block level including predefined groups from package mlr3 (method "cv"). A factor variable is used to define the grouping. Each class is either assigned to the test or training set.', out.width="40%", fig.pos="ht", eval=FALSE}
autoplot(rsmp_cv_group, size = 0.8, task = task_cv, fold_id = 1)
```

```{r cv-eval-1, echo=FALSE}
task_cv = tsk("ecuador")
group = as.factor(kmeans(task$coordinates(), 8)$cluster)
task_cv$cbind(data.frame("group" = group))
task_cv$set_col_roles("group", roles = "group")

rsmp_cv_group = rsmp("cv", folds = 3)$instantiate(task_cv)

print(rsmp_cv_group$instance)
```

```{r cv-eval-2, fig.cap='Cross-Validation at the block level including predefined groups from package mlr3 (method "cv"). A factor variable is used to define the grouping. Each class is either assigned to the test or training set.', out.width="40%", fig.pos="ht", echo=FALSE}
autoplot(rsmp_cv_group, size = 0.8, task = task_cv, fold_id = 1) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

plot_env_single = autoplot(rsmp_env, size = 0.5, fold_id = 1, task = task)

plot_env_multi = autoplot(rsmp_env_multi, size = 0.5, fold_id = 1, task = task)

@@

```{r env-1-eval, fig.cap='Environmental leave-one-block-out CV from package blockCV using one (left, "distdeforest") and two (right, "distdeforest" and "slope") predictors to define blocks in the feature space. Due to feature space clustering observations are not (necessarily) grouped in the spatial domain.', out.width="100%", fig.pos="ht", echo=FALSE}
rsmp_env = rsmp("spcv_env",
  features = "distdeforest", folds = 5)

@@

plot_env_single = autoplot(rsmp_env,
  size = 0.5, fold_id = 1, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02))

plot_env_multi = autoplot(rsmp_env_multi,
  size = 0.5, fold_id = 1, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02))

library("patchwork")
plot_env_single + plot_env_multi + plot_layout(guides = "collect")
```

@@

```{r llto, echo=FALSE, out.width="70%", fig.cap='Perspective plot of leave-location-and-time-out CV from package CAST (method "sptcv\\_cstf" and arguments time\\_var = "Date" and space\\_var = "SOURCEID"). The grey points are excluded from both the training and the test set in this fold.', fig.pos="h"}
knitr::include_graphics("png/llto.png")
```

@@

```{r create-task}
backend = mlr3::as_data_backend(ecuador)
task = TaskClassifST$new(
  id = "ecuador", backend = backend, target = "slides", positive = "TRUE",
  extra_args = list(
    coordinate_names = c("x", "y"), coords_as_features = FALSE,
    crs = "EPSG:32717")
)
```

@@

```{r rr-nspcv-1, cache.lazy = FALSE, results='hide'}
rsmp_nsp = rsmp("repeated_cv", folds = 4, repeats = 2)
rsmp_nsp
rr_nsp = resample(
  task = task, learner = learner,
  resampling = rsmp_nsp
)
```

@@

```{r rr-spcv-1, cache.lazy = FALSE, results='hide'}
rsmp_sp = rsmp("repeated_spcv_coords", folds = 4, repeats = 2)
rsmp_sp
rr_sp = resample(
  task = task, learner = learner,
  resampling = rsmp_sp
)
```
